---
title: "meta analysis for all defense traits"
author: "Suz"
date: "24 January 2021"
output: html_document
---

libraries required

```{r}
library(dplyr)
library(metafor)
library(glmulti)
library(ggplot2)
library(gridExtra)
```

Read in data

```{r}
defensesdata <- read.csv("defensesdatafull.csv")
defensesmetric <- read.csv("defensemetric.csv") %>%
  select(Species, PC1)

#add the metric for defense to the 
defensesdata <- left_join(defensesdata, defensesmetric, by = "Species")
```

Create create function for rma.glmulti- rma with yi and vi

```{r}
rma.glmulti <- function(formula, vi, data, ...) {
  rma(as.formula(paste(deparse(formula))), vi=vi, data=data, method = "ML", ...)
}
```

Also create a function for getting the moderator importance and significance averaged across all models
```{r}
setMethod('getfit', 'rma.uni', function(object, ...){
  if (object$test == "z") {
    cbind(estimate = coef(object), se = sqrt(diag(vcov(object))), df= 10000)
  } else {
    cbind(estimate = coef(object), se = sqrt(diag(vcov(object))), df= object$k - object$p)
  }
})
```

Run meta-analytic multi-model selection on each defense trait and the metric for all defenses against all climate variables

1. C:N
```{r}
C.N.modelselection <- glmulti(C.N.LRR ~ Changeavtemp + Changetemprange + Changetempvar + Changeav + Changerange + Changevar + Changemaxseasonal + Changeminseasonal + Changedrought + Changeheatwave + Changevpd + Changemaxdryspell + Changedrought, vi = 1, data=defensesdata, level=1, fitfunction=rma.glmulti, crit="aicc", confsetsize = 8192)

summary(C.N.modelselection@objects[[1]])
print(C.N.modelselection)
plot(C.N.modelselection)

mmi <- as.data.frame(coef(C.N.modelselection))
mmi <- data.frame(Estimate=mmi$Est, SE=sqrt(mmi$Uncond), Importance=mmi$Importance, row.names=row.names(mmi))
mmi$z <- mmi$Estimate / mmi$SE
mmi$p <- 2*pnorm(abs(mmi$z), lower.tail=FALSE)
names(mmi) <- c("Estimate", "Std. Error", "Importance", "z value", "Pr(>|z|)")
mmi$ci.lb <- mmi[[1]] - qnorm(.975) * mmi[[2]]
mmi$ci.ub <- mmi[[1]] + qnorm(.975) * mmi[[2]]
mmi <- mmi[order(mmi$Importance, decreasing=TRUE), c(1,2,4:7,3)]
round(mmi, 4)
```

2. Lipids

```{r}
lipidsmodelselection <- glmulti(lipidsLRR ~ Changeavtemp + Changetemprange + Changetempvar + Changeav + Changerange + Changevar + Changemaxseasonal + Changeminseasonal + Changedrought + Changeheatwave + Changevpd + Changemaxdryspell + Changedrought, vi = 1, data=defensesdata, level=1, fitfunction=rma.glmulti, crit="aicc", confsetsize = 8192)

summary(lipidsmodelselection@objects[[1]])
print(lipidsmodelselection)
plot(lipidsmodelselection)

mmi <- as.data.frame(coef(lipidsmodelselection))
mmi <- data.frame(Estimate=mmi$Est, SE=sqrt(mmi$Uncond), Importance=mmi$Importance, row.names=row.names(mmi))
mmi$z <- mmi$Estimate / mmi$SE
mmi$p <- 2*pnorm(abs(mmi$z), lower.tail=FALSE)
names(mmi) <- c("Estimate", "Std. Error", "Importance", "z value", "Pr(>|z|)")
mmi$ci.lb <- mmi[[1]] - qnorm(.975) * mmi[[2]]
mmi$ci.ub <- mmi[[1]] + qnorm(.975) * mmi[[2]]
mmi <- mmi[order(mmi$Importance, decreasing=TRUE), c(1,2,4:7,3)]
round(mmi, 4)
```

3. Ash

```{r}
ashmodelselection <- glmulti(ashLRR ~ Changeavtemp + Changetemprange + Changetempvar + Changeav + Changerange + Changevar + Changemaxseasonal + Changeminseasonal + Changedrought + Changeheatwave + Changevpd + Changemaxdryspell + Changedrought, vi = 1, data=defensesdata, level=1, fitfunction=rma.glmulti, crit="aicc", confsetsize = 8192)

summary(ashmodelselection@objects[[1]])
print(ashmodelselection)
plot(ashmodelselection)

mmi <- as.data.frame(coef(ashmodelselection))
mmi <- data.frame(Estimate=mmi$Est, SE=sqrt(mmi$Uncond), Importance=mmi$Importance, row.names=row.names(mmi))
mmi$z <- mmi$Estimate / mmi$SE
mmi$p <- 2*pnorm(abs(mmi$z), lower.tail=FALSE)
names(mmi) <- c("Estimate", "Std. Error", "Importance", "z value", "Pr(>|z|)")
mmi$ci.lb <- mmi[[1]] - qnorm(.975) * mmi[[2]]
mmi$ci.ub <- mmi[[1]] + qnorm(.975) * mmi[[2]]
mmi <- mmi[order(mmi$Importance, decreasing=TRUE), c(1,2,4:7,3)]
round(mmi, 4)
```

4. Phenolics

```{r}
phenolicsmodelselection <- glmulti(phenolicsLRR ~ Changeavtemp + Changetemprange + Changetempvar + Changeav + Changerange + Changevar + Changemaxseasonal + Changeminseasonal + Changedrought + Changeheatwave + Changevpd + Changemaxdryspell + Changedrought, vi = 1, data=defensesdata, level=1, fitfunction=rma.glmulti, crit="aicc", confsetsize = 8192)

summary(phenolicsmodelselection@objects[[1]])
print(phenolicsmodelselection)
plot(phenolicsmodelselection)

mmi <- as.data.frame(coef(phenolicsmodelselection))
mmi <- data.frame(Estimate=mmi$Est, SE=sqrt(mmi$Uncond), Importance=mmi$Importance, row.names=row.names(mmi))
mmi$z <- mmi$Estimate / mmi$SE
mmi$p <- 2*pnorm(abs(mmi$z), lower.tail=FALSE)
names(mmi) <- c("Estimate", "Std. Error", "Importance", "z value", "Pr(>|z|)")
mmi$ci.lb <- mmi[[1]] - qnorm(.975) * mmi[[2]]
mmi$ci.ub <- mmi[[1]] + qnorm(.975) * mmi[[2]]
mmi <- mmi[order(mmi$Importance, decreasing=TRUE), c(1,2,4:7,3)]
round(mmi, 4)
```

4. Defenses metric

```{r}
overallmodelselection <- glmulti(PC1 ~ Changeavtemp + Changetemprange + Changetempvar + Changeav + Changerange + Changevar + Changemaxseasonal + Changeminseasonal + Changedrought + Changeheatwave + Changevpd + Changemaxdryspell + Changedrought, vi = 1, data=defensesdata, level=1, fitfunction=rma.glmulti, crit="aicc", confsetsize = 8192)

summary(overallmodelselection@objects[[1]])
print(overallmodelselection)
plot(overallmodelselection)

mmi <- as.data.frame(coef(overallmodelselection))
mmi <- data.frame(Estimate=mmi$Est, SE=sqrt(mmi$Uncond), Importance=mmi$Importance, row.names=row.names(mmi))
mmi$z <- mmi$Estimate / mmi$SE
mmi$p <- 2*pnorm(abs(mmi$z), lower.tail=FALSE)
names(mmi) <- c("Estimate", "Std. Error", "Importance", "z value", "Pr(>|z|)")
mmi$ci.lb <- mmi[[1]] - qnorm(.975) * mmi[[2]]
mmi$ci.ub <- mmi[[1]] + qnorm(.975) * mmi[[2]]
mmi <- mmi[order(mmi$Importance, decreasing=TRUE), c(1,2,4:7,3)]
round(mmi, 4)
```
