---
title: "PCA defenses"
author: "Suz"
date: "7 January 2021"
output: html_document
---

```{r}
library(missMDA)
library(FactoMineR)
library(dplyr)
library(tidyr)
library(ggfortify)
library(ggplot2)
```

Reading in Data created previously in rmarkdown 
```{r}
defensedata <- read.csv("defensesdatafull.csv")
```

Create a dataframe with just the three variables to see where the NAs lie 

```{r}

defensePCAdata <- defensedata[c("Species", "Changeavtemp", "Changetempvar", "Changetemprange", "Changeav", "Changevar", "Changerange", "Changedrought", "Changemaxseasonal", "Changeminseasonal", "Changeheatwave", "Changevpd", "Changemaxdryspell", "phenolicsLRR", "lipidsLRR", "thicknessLRR", "C.N.LRR", "ashLRR")]

```

A few species have no data at all (due to lack of germination or death in glasshouse). These species we will remove as they are of no use in a PCA. A few remaining species will need some data imputation as they have a missing value here and there.

Create a wide data frame so that each species gets a PCA for the three variables


```{r}
#select the data of species and the five traits

defensesPCAdata1 <- defensePCAdata[c("Species", "phenolicsLRR", "lipidsLRR", "thicknessLRR", "C.N.LRR", "ashLRR")]

```

Now I want to remove the species that have NAs across all three leaf economic traits - these cannot be imputed and do not contribute any data

```{r}
defensesPCAdataspecies <- filter(defensesPCAdata1, Species != "Acacia aneura" & Species != "Acronychia imperforata" & Species != "Allocasuarina monilifera" & Species != "Austrostipa densiflora" & Species != "Callitris glaucophylla" & Species != "Crowea exalata" & Species != "Eucalyptus fastigata" & Species != "Eucalyptus langleyi" & Species != "Eucalyptus pauciflora" & Species != "Gompholobium latifolium" & Species != "Grevillea mucronulata" & Species != "Polycalymma stuartii" & Species != "Xerochrysum viscosum")

defensesPCAdatafinal <- defensesPCAdataspecies %>% select(2:6) ## selecting only the numeric columns
```

Run a PCA analysis on the defense trait variables 
```{r}
defensesPCAimputed <- imputePCA(defensesPCAdatafinal)

defenses.PCA <- prcomp(t(defensesPCAimputed$completeObs))

plot(defenses.PCA)
summary(defenses.PCA)
str(defenses.PCA)
```

Save that PCA analysis as a dataframe and bind with the species names and data 

```{r}
defensemetric1 <- cbind(defensesPCAdataspecies, defenses.PCA$rotation) #for some reason the function merge adds a tgonne of duplicated rows so now I remove the duplicated rows:
defensemetric <- left_join(defensemetric1, defensePCAdata, by = "Species")

#and now I wanna remove the species that don't have the PCA data (didn't have any data)

defensemetric <- filter(defensemetric, Species != "Acacia aneura" & Species != "Acronychia imperforata" & Species != "Allocasuarina monilifera" & Species != "Austrostipa densiflora" & Species != "Callitris glaucophylla" & Species != "Crowea exalata" & Species != "Eucalyptus fastigata" & Species != "Eucalyptus langleyi" & Species != "Eucalyptus pauciflora" & Species != "Gompholobium latifolium" & Species != "Grevillea mucronulata" & Species != "Polycalymma stuartii" & Species != "Xerochrysum viscosum")

```

save that data as a csv to use in further analysis

```{r}
write.csv(defensemetric, "defensemetric.csv")
```

