---
title: "testing phylogenetic signal in defence trait changes"
output: html_document
date: "2025-05-29"
---

Libraries required
```{r}
library(tidyverse)
library(ggtree)
library(viridis)
library(phylobase)
library(phylosignal)
library(readr)
library(phytools)
```
Read in my data
```{r}
defencedata <- read.csv("defensesdatafull.csv") %>%
        mutate(Species=as.character(Species)) %>%
        filter(Species!="Acacia hakeoidesT") %>%
        mutate(Species = ifelse(Species=="Acacia hakeoidesB", "Acacia hakeoides", Species))

defencemetric <- read.csv("defensemetric.csv") %>%
        mutate(Species=as.character(Species)) %>%
        filter(Species!="Acacia hakeoidesT") %>%
        mutate(Species = ifelse(Species=="Acacia hakeoidesB", "Acacia hakeoides", Species)) %>%
        select(Species, defenseoverallLRR)

defencedata <- left_join(defencedata, defencemetric, by = "Species")
```

Read in the phylogenetic tree
```{r}
bigtree <- read.tree(file = "PhytoPhylo.tre") #reading tree
specieslist <- read.csv("specieslist.csv", header = T) #reading species list
nodes <- read_table("nodes.csv") #reading nodes - nodes were downloaded from the S.phylomaker github page
source("https://raw.githubusercontent.com/jinyizju/S.PhyloMaker/master/R_codes%20for%20S.PhyloMaker")
resultphylo <- S.PhyloMaker(spList = specieslist, tree = bigtree, nodes = nodes, scenarios = "S3") #pruning tree
str(resultphylo) #looking at results of my tree
```

Plot the tree
```{r}
#plotting tree
tiff("mytree.tiff", width = 6, height = 8, units = 'in', res = 300)
plot(resultphylo$Scenario.3,main="Scenarion Three") ## plotting tree
dev.off()
```

Prune the tree and add it to the defence data
```{r}
PrunedTree <- resultphylo$Scenario.3 ## this is my pruned tree, i picked scen 3 because it deals with missing spp best

Species_ <- gsub(" ", "_", defencedata$Species)
defencedata <- cbind(defencedata, Species_)
defencedata <- rename(defencedata, "PrunedTree$tip.label"=Species_)
defencedata <- right_join(defencedata, as.data.frame(PrunedTree$tip.label), by = "PrunedTree$tip.label")
defencedata <- rename(defencedata, Species_="PrunedTree$tip.label")
species.vector <- as.vector(defencedata$Species_)

#adds a column "phylo" to your data 
defencedata[, "phylo"] <- species.vector
branchlengths <- compute.brlen(PrunedTree)
```

Run phylogenetic analysis
```{r}
phylo4d <- phylo4d(branchlengths, defencedata$phenolicsLRR) ### in here and in the next two test rows below, 
                                    #Change the "phenolicsLRR" variable to each remaining trait (lipids, C:N etc) until you have analysed all the trait variables, 
                                            
                                    #note that the variables are:
                                            # phenolicsLRR
                                            # lipidsLRR
                                            # thicknessLRR
                                            # C.N.LRR
                                            # ashLRR
                                            # lmaLRR
                                            # defenseoverallLRR
                                            

teststat <- phyloSignal(phylo4d, rep =999)

teststat

```

Two methods for getting signal (more in phylosig documentation)
```{r}
#Method for signal 1 (K)
phylosig(branchlengths, (defencedata$defenseoverallLRR), method = "K", test = T)

#Method for signal 2 (lamba)
phylosig(branchlengths, (defencedata$defenseoverallLRR), method = "lambda", test = T)

## and keep replacing afer the "$" for each variable as with the previous chunk
```

Make a bar plot of some sort that matches the same tree order to put in the supplementary to show heat map of defense trait change on phylogeny 

Having to do this mostly by hand as I've found it's the easiest way
```{r}
# 1. Create a species list from the tree
specieslist <- as.data.frame(resultphylo$Scenario.3$tip.label) %>%
  rename(Species = `resultphylo$Scenario.3$tip.label`)

specieslist$Species <- gsub("_", " ", specieslist$Species)

# 2. merge with defence dataa
defencewithspecieslist <- left_join(specieslist, defencedata, by = "Species")

# 3. lock in order of species
defencewithspecieslist$Species <- factor(defencewithspecieslist$Species, levels = defencewithspecieslist$Species)

# 4. plot

defenceyphyloplot <- ggplot(defencewithspecieslist, aes(x=defenseoverallLRR, y=Species)) +
  geom_col(fill = "darkseagreen4") +
  theme_classic() +
  theme(axis.text.y = element_text(face = "italic", hjust = 0)) +
  theme(axis.text.y=element_text(size=10)) +
  theme(legend.text = element_blank())
plot(defenceyphyloplot)

# 5. save the plot and stitch it together with the tree in an image editor
ggsave(filename = "phyloplotstrength.tiff", plot=defenceyphyloplot, device="tiff", width=5, height=8, units="in")
```

