---
title: "two way defense and means analysis"
author: "Suz"
date: "10 November 2020"
output: html_document
---

Libraries required

```{r}
library(ggplot2)
library(metafor)
library(dplyr)
library(gridExtra)
```

Read in main data sheet for analyses

```{r}
defensedata <- read.csv("defensesdatafull.csv")
```

1. C:N vs mean temperature

```{r}
C.N.temp <- rma(yi=C.N.LRR, vi = C.N.vi, mods = Changeavtemp, data = defensedata)
```
```{r}
print(C.N.temp)
```

2. C:N to mean precipitation

```{r}
C.N.precipt <- rma(yi=C.N.LRR, vi = C.N.vi, mods = Changeav, data = defensedata)
```
```{r}
print(C.N.precipt)
```

3. Ash vs mean temperature

```{r}
ash.temp <- rma(yi=ashLRR, vi = ash.vi, mods = Changeavtemp, data = defensedata)
```
```{r}
print(ash.temp)
```

4. Ash vs. mean precip

```{r}
ashprecipt <- rma(yi=ashLRR, vi = ash.vi, mods = Changeav, data = defensedata)
```
```{r}
print(ashprecipt)
```

5. Lipids vs mean temp

```{r}
lipids.temp <- rma(yi=lipidsLRR, vi = lipids.vi, mods = Changeavtemp, data = defensedata)
```
```{r}
print(lipids.temp)
```
6. Lipids vs mean precip

```{r}
lipidsprecipt <- rma(yi=lipidsLRR, vi = lipids.vi, mods = Changeav, data = defensedata)
```
```{r}
print(lipidsprecipt)
```

7. Phenolics vs mean temp

```{r}
phenolics.temp <- rma(yi=phenolicsLRR, vi = phenolics.vi, mods = Changeavtemp, data = defensedata)
```
```{r}
print(phenolics.temp)
```

8. Phenolics vs mean precip

```{r}
phenolicsprecipt <- rma(yi=phenolicsLRR, vi = phenolics.vi, mods = Changeav, data = defensedata)
```
```{r}
print(phenolicsprecipt)
```

Plotting relationships for figure 1
- Each relationship is plotted one by one between mean temperature and the leaf trait change metric and mean precipitation and the leaf trait change metric
- Colour scheme for manuscripts is lightsalmon3 and darkseagreen4
- combine all plots together
- save tiff

1. Plots for both relationships of C:N

```{r}
C.N.tempplot <- ggplot(defensedata, aes(Changeavtemp, C.N.LRR)) + 
  geom_point(colour="darkseagreen4", size=4) +
  labs(cex=15, x="", y="Change in C:N") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.y = element_text(size=18)) +
  theme(axis.text.x = element_blank())
plot(C.N.tempplot)

C.N.precipplot <- ggplot(defensedata, aes(Changeav, C.N.LRR)) + ##Changeav is change in average precipitation
  geom_point(colour="darkseagreen4", size=4) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=8, x="", y="") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_blank())
plot(C.N.precipplot)
```

2. Plots for both relationships for Ash

```{r}
ashtempplot <- ggplot(defensedata, aes(Changeavtemp, ashLRR)) + 
  geom_point(colour="darkseagreen4", size=4) +
  labs(cex=15, x="", y="Change in Ash Content") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.y = element_text(size=18)) +
  theme(axis.text.x = element_blank())
plot(ashtempplot)

ashprecipplot <- ggplot(defensedata, aes(Changeav, ashLRR)) + ##Changeav is change in average precipitation
  geom_point(colour="darkseagreen4", size=4) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=8, x="", y="") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_blank())
plot(C.N.precipplot)
```

3. Both relationships for Lipids

```{r}
lipidstempplot <- ggplot(defensedata, aes(Changeavtemp, lipidsLRR)) + 
  geom_point(colour="darkseagreen4", size=4) + 
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=15, x="", y="Change in Lipids") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.y = element_text(size=18)) +
  theme(axis.text.x = element_blank())
plot(lipidstempplot)

lipidsprecipplot <- ggplot(defensedata, aes(Changeav, lipidsLRR)) + ##Changeav is change in average precipitation
  geom_point(colour="darkseagreen4", size=4) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=8, x="", y="") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_blank())
plot(lipidsprecipplot)
```

4. Both plots for relationships of phenolics

```{r}
phenolicstempplot <- ggplot(defensedata, aes(Changeavtemp, phenolicsLRR)) + 
  geom_point(colour="darkseagreen4", size=4) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=15, x="Change in mean temperature", y="Change in Phenolics") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.y = element_text(size=18)) +
  theme(axis.text.x = element_text(size=18))
plot(phenolicstempplot)

phenolicsprecipplot <- ggplot(defensedata, aes(Changeav, phenolicsLRR)) + ##Changeav is change in average precipitation
  geom_point(colour="darkseagreen4", size=4) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=8, x="Change in mean precipitation", y="") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(axis.text.x=element_text(size=18)) +
  theme(axis.text.y =element_blank()) +
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5))
plot(phenolicsprecipplot)
```

Stitch all plots together 

```{r}
figure1 <- grid.arrange(C.N.tempplot, C.N.precipplot,
                        ashtempplot, ashprecipplot,
                        lipidstempplot, lipidsprecipplot,
                        phenolicstempplot, phenolicsprecipplot, ncol=2)
```

And save the plot

```{r}
ggsave(
  "figure2.tiff",
  plot=figure1,
  device = "tiff",
  width=12,
  height=16,
  dpi=200
)
```


