---
title: "two way defense and means analysis"
author: "Suz"
date: "10 November 2020"
output: html_document
---

Libraries required

```{r}
library(ggplot2)
library(metafor)
library(dplyr)
library(gridExtra)
```

Read in main data sheet for analyses

```{r}
defensedata <- read.csv("defensesdatafull.csv")
```

First compare changes through time in the defence traits (absolute change between historic and modern populations)

```{r}
#reshape data to visualise the differences in trait changes through time
defensecahngethroughtimedata <- defensedata %>%
  select(Species, thicknessLRR, C.N.LRR, ashLRR, lipidsLRR, phenolicsLRR) %>%
  pivot_longer(!Species, names_to = "trait", values_to = "LRR") %>%
  mutate(absolutetraitchange = abs(LRR))

## plot a figure with these differences in trait changes
defencetraitchangeplot <- ggplot(defensecahngethroughtimedata, aes(x=trait, y = absolutetraitchange)) +
  geom_quasirandom(color = "darkseagreen4", size = 4) +
  stat_summary(
    fun = mean,
    geom = "point",
    shape = 15,         # Square shape
    size = 6,
    color = "lightsalmon3"
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.4,
    color = "lightsalmon3"
  ) +
  theme_classic(base_size  = 20)

defencetraitchangeplot

## save this plot
ggsave(filename = "traitchangethroughtime.tiff", plot=defencetraitchangeplot, device="tiff", dpi=100, width=15, height=10, units="in")
```

Then look at correlations between climate of the seed collection and the trait
```{r}
#Some hectic dataframe wrangling to get this where it needs to be
defenseclimatespatialpatternssuffix <- defensedata %>%
  mutate(Olddryspell = pmax(Old.max.Ndays.dry.spell.y1...., Old.max.Ndays.dry.spell.y2...., Old.max.Ndays.dry.spell.y3...., Old.max.Ndays.dry.spell.y4...., Old.max.Ndays.dry.spell.y5....)) %>%
  mutate(Moddryspell = pmax(Modern.max.Ndays.dry.spell.y1...., Modern.max.Ndays.dry.spell.y2...., Modern.max.Ndays.dry.spell.y3...., Modern.max.Ndays.dry.spell.y4...., Modern.max.Ndays.dry.spell.y5....)) %>%
  select(Species, Oldavtemp, Oldtempvar, Oldtemprange, Modavtemp, Modtempvar, Modtemprange, Oldmonth, Oldvar, Oldrange, Oldmaxdrought, Oldaverageseasonalprecipmax, Oldaverageseasonalprecipmin, Modmonth, Modvar, Modrange, Modmaxdrought, Modernprecipseasonalmaxaverage, Modernprecipminseasonalaverage, Oldheatwave, Modheatwave, Olddryspell, Moddryspell, Oldavvpd, Modavtemp, Initialphenolics_Mod, Initialphenolics_Old, Lipids.mg._Mod, Lipids.mg._Old, avg.thickness_Mod, avg.thickness_Old, C.N_Mod, C.N_Old, Ash_Mod, Ash_Old) %>%
  pivot_longer(
    cols = matches("(_|\\.)?(Old|Mod)$"),
    names_to = c("variable", "source"),
    names_pattern = "^(.*)[._](Old|Mod)$"
  ) %>%
  select(Species, variable, source, value)


defenseclimatespatialpatternsprefix <- defensedata %>%
  mutate(Olddryspell = pmax(Old.max.Ndays.dry.spell.y1...., Old.max.Ndays.dry.spell.y2...., Old.max.Ndays.dry.spell.y3...., Old.max.Ndays.dry.spell.y4...., Old.max.Ndays.dry.spell.y5....)) %>%
  mutate(Moddryspell = pmax(Modern.max.Ndays.dry.spell.y1...., Modern.max.Ndays.dry.spell.y2...., Modern.max.Ndays.dry.spell.y3...., Modern.max.Ndays.dry.spell.y4...., Modern.max.Ndays.dry.spell.y5....)) %>%
  select(Species, Oldavtemp, Oldtempvar, Oldtemprange, Modavtemp, Modtempvar, Modtemprange, Oldmonth, Oldvar, Oldrange, Oldmaxdrought, Oldaverageseasonalprecipmax, Oldaverageseasonalprecipmin, Modmonth, Modvar, Modrange, Modmaxdrought, Modernprecipseasonalmaxaverage, Modernprecipminseasonalaverage, Oldheatwave, Modheatwave, Olddryspell, Moddryspell, Oldavvpd, Modavtemp, Initialphenolics_Mod, Initialphenolics_Old, Lipids.mg._Mod, Lipids.mg._Old, avg.thickness_Mod, avg.thickness_Old, C.N_Mod, C.N_Old, Ash_Mod, Ash_Old) %>%
  pivot_longer(
    cols = matches("^(Old|Mod)"),
    names_to = c("source", "variable"),
    names_pattern = "^(Old|Mod)(.*)"
  ) %>%
  select(Species, variable, source, value)

defenseclimatespatialpatterns <- bind_rows(defenseclimatespatialpatternssuffix, defenseclimatespatialpatternsprefix) %>%
  pivot_wider(
    names_from = variable,
    values_from = value,
    id_cols = c(Species, source)
  ) %>%
  mutate(avg.thickness = log10(avg.thickness))

#now run linear regressions on the defence trait versus climate variables
response_vars <- c("Initialphenolics", "Lipids.mg.", "avg.thickness", "C.N", "Ash")
predictor_vars <- c("avtemp", "tempvar", "temprange", "month", "var",
                    "range", "maxdrought", "averageseasonalprecipmax", "averageseasonalprecipmin", "heatwave",
                    "dryspell", "avvpd")

# Run linear regressions for each response x predictor variable combination
results <- map_dfr(response_vars, function(response) {
  map_dfr(predictor_vars, function(predictor) {
    # Perform the regression
    model <- lm(as.formula(paste(response, "~", predictor)), data = defenseclimatespatialpatterns)
    
    # Tidy the results to extract model coefficients and p-values
    tidy_model <- tidy(model)
    glance_model <- glance(model)
    
    # Add the response and predictor variables to the output
    tidy_model %>%
      mutate(response = response, predictor = predictor, r.squared = glance_model$r.squared)
  })
})

# View the results
print(results)

#save the results
write.csv(results, file = "spatialclimateregressions.csv", row.names = FALSE)
```

Create a figure for the mean temp and precip versus each defense trait (10 panel figure)
```{r}
#C:N
C.N.tempplot <- ggplot(defenseclimatespatialpatterns, aes(avtemp, C.N)) + 
  geom_point(colour="darkseagreen4", size=4) +
  labs(cex=15, x="", y="C:N") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_text(size=18))
plot(C.N.tempplot)

C.N.precipplot <- ggplot(defenseclimatespatialpatterns, aes(month, C.N)) + 
  geom_point(colour="darkseagreen4", size=4) +
  labs(cex=15, x="", y="C:N") +
  theme_bw() +
  scale_x_continuous(limits = c(0.8, 4.5)) +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_text(size=18))
plot(C.N.precipplot)

#Ash
Ash.tempplot <- ggplot(defenseclimatespatialpatterns, aes(avtemp, Ash)) + 
  geom_point(colour="darkseagreen4", size=4) +
  labs(cex=15, x="", y="ash") +
  geom_smooth(method = "lm", colour = "lightsalmon3", se = F) +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_text(size=18))
plot(Ash.tempplot)

Ash.precipplot <- ggplot(defenseclimatespatialpatterns, aes(month, Ash)) + 
  geom_point(colour="darkseagreen4", size=4) +
  labs(cex=15, x="", y="ash") +
  theme_bw() +
  scale_x_continuous(limits = c(0.8, 4.5)) +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_text(size=18))
plot(Ash.precipplot)

#Lipids
Lipids.tempplot <- ggplot(defenseclimatespatialpatterns, aes(avtemp, Lipids.mg.)) + 
  geom_point(colour="darkseagreen4", size=4) +
  labs(cex=15, x="", y="lipids") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_text(size=18))
plot(Lipids.tempplot)

Lipids.precipplot <- ggplot(defenseclimatespatialpatterns, aes(month, Lipids.mg.)) + 
  geom_point(colour="darkseagreen4", size=4) +
  labs(cex=15, x="", y="lipids") +
  theme_bw() +
  scale_x_continuous(limits = c(0.8, 4.5)) +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_text(size=18))
plot(Lipids.precipplot)

#phenolics
Phenolics.tempplot <- ggplot(defenseclimatespatialpatterns, aes(avtemp, Initialphenolics)) + 
  geom_point(colour="darkseagreen4", size=4) +
  labs(cex=15, x="", y="phenolics") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_text(size=18))
plot(Phenolics.tempplot)

Phenolics.precipplot <- ggplot(defenseclimatespatialpatterns, aes(month, Initialphenolics)) + 
  geom_point(colour="darkseagreen4", size=4) +
  geom_smooth(method = "lm", colour = "lightsalmon3", se = F) +
  labs(cex=15, x="", y="phenolics") +
  theme_bw() +
  scale_x_continuous(limits = c(0.8, 4.5)) +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_text(size=18))
plot(Phenolics.precipplot)

#leaf thickness
thickness.tempplot <- ggplot(defenseclimatespatialpatterns, aes(avtemp, avg.thickness)) + 
  geom_point(colour="darkseagreen4", size=4) +
  labs(cex=15, x="", y="thickness") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_text(size=18))
plot(thickness.tempplot)

thickness.precipplot <- ggplot(defenseclimatespatialpatterns, aes(month, avg.thickness)) + 
  geom_point(colour="darkseagreen4", size=4) +
  labs(cex=15, x="", y="thickness") +
  theme_bw() +
  scale_x_continuous(limits = c(0.8, 4.5)) +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_text(size=18))
plot(thickness.precipplot)

#save all these
ggsave(filename = "CNtempplot.tiff", plot=C.N.tempplot, device="tiff", width=8, height=5, units="in")
ggsave(filename = "CNprecipplot.tiff", plot=C.N.precipplot, device="tiff", width=8, height=5, units="in")
ggsave(filename = "ashtempplot.tiff", plot=Ash.tempplot, device="tiff", width=8, height=5, units="in")
ggsave(filename = "ashprecipplot.tiff", plot=Ash.precipplot, device="tiff", width=8, height=5, units="in")
ggsave(filename = "lipidstempplot.tiff", plot=Lipids.tempplot, device="tiff", width=8, height=5, units="in")
ggsave(filename = "lipidsprecipplot.tiff", plot=Lipids.precipplot, device="tiff", width=8, height=5, units="in")
ggsave(filename = "phenolicstempplot.tiff", plot=Phenolics.tempplot, device="tiff", width=8, height=5, units="in")
ggsave(filename = "phenolicsprecipplot.tiff", plot=Phenolics.precipplot, device="tiff", width=8, height=5, units="in")
ggsave(filename = "thicknesstempplot.tiff", plot=thickness.tempplot, device="tiff", width=8, height=5, units="in")
ggsave(filename = "thicknessprecipplot.tiff", plot=thickness.precipplot, device="tiff", width=8, height=5, units="in")

```

Now analyse the actual hypothesised relationships between mean temperature/precipitation CHANGE through time and Trait CHANGE through time

1. C:N vs mean temperature

```{r}
C.N.temp <- rma(yi=C.N.LRR, vi = C.N.vi, mods = Changeavtemp, data = defensedata)
```
```{r}
print(C.N.temp)
```

2. C:N to mean precipitation

```{r}
C.N.precipt <- rma(yi=C.N.LRR, vi = C.N.vi, mods = Changeav, data = defensedata)
```
```{r}
print(C.N.precipt)
```

3. Ash vs mean temperature

```{r}
ash.temp <- rma(yi=ashLRR, vi = ash.vi, mods = Changeavtemp, data = defensedata)
```
```{r}
print(ash.temp)
```

4. Ash vs. mean precip

```{r}
ashprecipt <- rma(yi=ashLRR, vi = ash.vi, mods = Changeav, data = defensedata)
```
```{r}
print(ashprecipt)
```

5. Lipids vs mean temp

```{r}
lipids.temp <- rma(yi=lipidsLRR, vi = lipids.vi, mods = Changeavtemp, data = defensedata)
```
```{r}
print(lipids.temp)
```
6. Lipids vs mean precip

```{r}
lipidsprecipt <- rma(yi=lipidsLRR, vi = lipids.vi, mods = Changeav, data = defensedata)
```
```{r}
print(lipidsprecipt)
```

7. Phenolics vs mean temp

```{r}
phenolics.temp <- rma(yi=phenolicsLRR, vi = phenolics.vi, mods = Changeavtemp, data = defensedata)
```
```{r}
print(phenolics.temp)
```

8. Phenolics vs mean precip

```{r}
phenolicsprecipt <- rma(yi=phenolicsLRR, vi = phenolics.vi, mods = Changeav, data = defensedata)
```
```{r}
print(phenolicsprecipt)
```

Plotting relationships for figure 1
- Each relationship is plotted one by one between mean temperature and the leaf trait change metric and mean precipitation and the leaf trait change metric
- Colour scheme for manuscripts is lightsalmon3 and darkseagreen4
- combine all plots together
- save tiff

1. Plots for both relationships of C:N

```{r}
C.N.tempplot <- ggplot(defensedata, aes(Changeavtemp, C.N.LRR)) + 
  geom_point(colour="darkseagreen4", size=4) +
  labs(cex=15, x="", y="Change in C:N") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.y = element_text(size=18)) +
  theme(axis.text.x = element_blank())
plot(C.N.tempplot)

C.N.precipplot <- ggplot(defensedata, aes(Changeav, C.N.LRR)) + ##Changeav is change in average precipitation
  geom_point(colour="darkseagreen4", size=4) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=8, x="", y="") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_blank())
plot(C.N.precipplot)
```

2. Plots for both relationships for Ash

```{r}
ashtempplot <- ggplot(defensedata, aes(Changeavtemp, ashLRR)) + 
  geom_point(colour="darkseagreen4", size=4) +
  labs(cex=15, x="", y="Change in Ash Content") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.y = element_text(size=18)) +
  theme(axis.text.x = element_blank())
plot(ashtempplot)

ashprecipplot <- ggplot(defensedata, aes(Changeav, ashLRR)) + ##Changeav is change in average precipitation
  geom_point(colour="darkseagreen4", size=4) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=8, x="", y="") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_blank())
plot(C.N.precipplot)
```

3. Both relationships for Lipids

```{r}
lipidstempplot <- ggplot(defensedata, aes(Changeavtemp, lipidsLRR)) + 
  geom_point(colour="darkseagreen4", size=4) + 
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=15, x="", y="Change in Lipids") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.y = element_text(size=18)) +
  theme(axis.text.x = element_blank())
plot(lipidstempplot)

lipidsprecipplot <- ggplot(defensedata, aes(Changeav, lipidsLRR)) + ##Changeav is change in average precipitation
  geom_point(colour="darkseagreen4", size=4) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=8, x="", y="") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_blank())
plot(lipidsprecipplot)
```

4. Both plots for relationships of phenolics

```{r}
phenolicstempplot <- ggplot(defensedata, aes(Changeavtemp, phenolicsLRR)) + 
  geom_point(colour="darkseagreen4", size=4) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=15, x="Change in mean temperature", y="Change in Phenolics") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.y = element_text(size=18)) +
  theme(axis.text.x = element_text(size=18))
plot(phenolicstempplot)

phenolicsprecipplot <- ggplot(defensedata, aes(Changeav, phenolicsLRR)) + ##Changeav is change in average precipitation
  geom_point(colour="darkseagreen4", size=4) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=8, x="Change in mean precipitation", y="") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(axis.text.x=element_text(size=18)) +
  theme(axis.text.y =element_blank()) +
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5))
plot(phenolicsprecipplot)
```

Stitch all plots together 

```{r}
figure1 <- grid.arrange(C.N.tempplot, C.N.precipplot,
                        ashtempplot, ashprecipplot,
                        lipidstempplot, lipidsprecipplot,
                        phenolicstempplot, phenolicsprecipplot, ncol=2)
```

And save the plot

```{r}
ggsave(
  "figure2.tiff",
  plot=figure1,
  device = "tiff",
  width=12,
  height=16,
  dpi=200
)
```


